Dark energy is due to a math error from 1935

### 2024-12-09 paper/paper.tex
Start with a re-write.

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{amsmath}
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{doi}
\usepackage{csquotes}

\def \thetitle {Correcting supernova luminosity for time dilation implies no dark energy}
\title{\thetitle}

@@

%%% Add PDF metadata to help others organize their library
%%% Once the PDF is generated, you can check the metadata with
%%% $ pdfinfo template.pdf
\hypersetup{
pdftitle={\thetitle},
pdfsubject={astro-ph.CO},
pdfauthor={Logan P.~Evans},
pdfkeywords={cosmological parameters, dark energy},
}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{lemma}{Lemma}

@@

\begin{abstract}
  Existing analysis of Type Ia supernova relies on the assumption that
  luminosity is not affected by time dilation. However, when luminosity is
  corrected for time dilation, the relationship between luminosity distance and
  redshift for Type Ia supernova becomes linear. This indicates that the
  expansion rate of the universe is not accelerating and there is no need for
  dark energy to explain observational data.
\end{abstract}

@@

TODO.

@@

The magnitude measurements for Type Ia supernova goes through a long chain of
data processing. The data used here was collected by the Dark Energy Survey
Collaboration, as summarized in \citet{abbott2024}, while \citet{vincenzi2024}
describes the data processing pipeline in more detail. However, the data
processing does not explicitly correct for either redshift or time dilation.

@@

\begin{figure}[h!]
  \includegraphics[width=\linewidth]{../graphs/mu_distance_vs_redshift.png}
  \caption{The relationship between distance and redshift for two treatments of
  magnitude data. The displayed points are roughly a third of the values in the
  full DES dataset, selected evenly to aid visibility. The $k(z) = 1$ treatment
  is clearly non-linear while the $k(z) = 1 + z$ treatment appears to be
  linear.}
  \label{fig:mu_distance_vs_redshift}
\end{figure}

>>>>

\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{amsmath}
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{csquotes}

\title{Dark energy is due to a math error from 1935}

@@

\hypersetup{
pdftitle={Dark energy is due to a math error from 1935},
pdfsubject={astro-ph.CO},
pdfauthor={Logan P.~Evans},
pdfkeywords={cosmological parameters, dark energy},
}

@@

\begin{abstract}
  Observations of distant objects, such as supernovae and galaxies, is dimmed
  due to three phenomena associated with redshift: energy levels per photon,
  spectral bandwidth, and time dilation. Observed magnitudes are reported with
  k-corrections that should correct for these dimming effects, but since 1935,
  the equations used to perform these k-corrections have been missing the
  correction factor for time dilation. When observations of Type Ia supernova
  are corrected for this missing factor, the relationship between luminosity
  distance and redshift for Type Ia supernova becomes linear. This indicates
  that the expansion rate of the universe is not accelerating and there is no
  need for dark energy to explain observational data.
\end{abstract}

@@

Observations of Type Ia supernova are fundamental to the study of cosmology.
These measurements are necessary to calibrate models that measure redshift and
estimate distance. This relationship, often called the Hubble-Lema\^{i}tre law,
[TODO: find citation] describes how quickly the universe is expanding. However,
\citet{riess1998} and \citet{perlmutter1999} presented evidence that there
isn't a linear relationship between redshift and distance, but instead, distant
objects are father away than their redshift would predict (see Figure
\ref{fig:mu_distance_vs_redshift}). This phenomenon, whatever its source, is
referred to as dark energy.

Type Ia supernova are used to explore the relationship between distances and
redshifts because these events always happen in similar ways, so the absolute
brightness is roughly always the same. An analogy is to imagine someone walking
in the dark and lighting matches. As long as we know how brightly a match burns
at a known distance, we can estimate the distance to any match by measuring the
appearant brightness before applying some geometry.

It turns out that measuring the apperant brightness of a Type Ia supernova is
non-trivial. The aspect of the process that we will explore here concerns how
redshift affects the light we observe. This problem is often referred to as
k-corrections, and one of the first mathematical treatments of the problem was
performed by \citet{desitter1934}.

Willem de Sitter recognized three issues that would reduce the observed
magnitude of a distant observation. First, a redshifted photon carries less
energy than when it was originally emitted. Second, a narrow spectrum of
emitted light will be spread out over a larger observed spectrum. Finally, time
dilation will reduce the rate at which an observer detects photons. The
correction for each of these issues is identical: take a measurement and
multiply it by the factor $1 + z$.

A year later, \citet{hubble1935} published a similar set of calculations for
k-corrections, but these equations used $(1 + z)^2$ instead of the $(1 + z)^3$
correction term used by de Sitter. The Hubble k-corrections have been used ever
since.

Hubble and de Sitter were clearly aware of all three dimming effects associated
with redshift, but by \citet{oke1968energy}, the two factors of $(1 + z)$ were
attributed to the change in energy and to the spectral bandwidth elongation,
which leaves time dilation as the factor that was omitted.

In 1935, the correction of $(1 + z)^2$ empirically lined up the data so that
there seemed to be a linear relationship between redshift and distance. It
wasn't until \citet{riess1998}, sixty-three years later, that the omission of a
third $(1 + z)$ correction factor resulted in faulty conclusions.

Those faulty conclusions, however, have been the basis for a substantial amount
of scientific interest. In 2011, three physicists received the Nobel Prize in
Physics for discovering that the expansion rate of the universe is accelerating
\citet{straumann2012}. The papers \citet{riess1998} and \citet{perlmutter1999}
have each received over 20000 citations. It's not clear how much cosmology work
will need to be revisited with corrected distance measurements, but it is a
non-trivial amount.

@@

The magnitude measurements for Type Ia supernova goes through a long chain of
data processing. The data used here was collected by the Dark Energy Survey
Collaboration, as summarized in \citet{abbott2024} and described in more detail
by \citet{vincenzi2024}. However, the data processing does not explicitly
correct for either redshift or time dilation.

@@

\begin{figure}[h!]
  \includegraphics[width=\linewidth]{mu_distance_vs_redshift.png}
  \caption{The relationship between distance and redshift for two treatments of
  magnitude data. The displayed points are roughly a third of the values in the
  full DES dataset, selected evenly to aid visibility. The $k(z) = 1$ treatment
  is clearly non-linear while the $k(z) = 1 + z$ treatment appears to be
  linear.}
  \label{fig:mu_distance_vs_redshift}
\end{figure}

\begin{figure}[h!]
  \includegraphics[width=\linewidth]{k-corrections_for_photon_counts.png}
  \caption{An example of how k-corrections take an observed spectrum and
  produce a rest-frame spectrum. [TODO]
  }
  \label{fig:mu_distance_vs_redshift}
\end{figure}

\begin{figure}[h!]
  \includegraphics[width=\linewidth]{velocity_vs_distance.png}
  \caption{The relationship between expansion velocity and distance. The slope of this graph demonstrates the Hubble constant.
  }
  \label{fig:mu_distance_vs_redshift}
\end{figure}

\begin{figure}[h!]
  \includegraphics[width=\linewidth]{bootstrapped_H0.png}
  \caption{A histogram of 100000 bootstrap trials measuring the Hubble constant
  $H0$. Each trial samples an absolute Type Ia magnitude $M \sim Norm(-19.2334,
  0.0404)$ based on data published by \citet{camarena2020}. It then samples,
  with replacement, a population of supernovae from the dataset published by
  \citet{abbott2024}. Finally, it uses the non-parametric linear regression
  technique described by \citet{siegel1982}. The result of the bootstrap is $H0
  \sim Norm(65.94, 1.29)$.
  }
  \label{fig:mu_distance_vs_redshift}
\end{figure}
